---
title: "Assignment 3: Data Exploration"
author: "Yosia Theo Napitupulu"
output: pdf_document
geometry: margin=2.54cm
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Exploration.

## Directions

1.  Rename this file `<FirstLast>_A03_DataExploration.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai.

The completed exercise is due on Sept 30th.

## Set up your R session

1.  Check your working directory, load necessary packages (tidyverse), and upload two datasets: the ECOTOX neonicotinoid dataset (ECOTOX_Neonicotinoids_Insects_raw.csv) and the Niwot Ridge NEON dataset for litter and woody debris (NEON_NIWO_Litter_massdata_2018-08_raw.csv). Name these datasets "Neonics" and "Litter", respectively. Be sure to include the subcommand to read strings in as factors.

```{r,  message = FALSE}
# 1. Set up working directory
setwd("E:/ENV872/EDA-Fall2022/")
getwd()

# 2. Load packages
library(tidyverse)

# 3. Import datasets
Neonics.data <- read.csv("./Data/Raw/ECOTOX_Neonicotinoids_Insects_raw.csv", stringsAsFactors = TRUE)
Litter.data <- read.csv("./Data/Raw/NEON_NIWO_Litter_massdata_2018-08_raw.csv", stringsAsFactors = TRUE)

#View(Neonics and Litter)
View(Neonics.data)
View(Litter.data)


```

## Learn about your system

2.  The neonicotinoid dataset was collected from the Environmental Protection Agency's ECOTOX Knowledgebase, a database for ecotoxicology research. Neonicotinoids are a class of insecticides used widely in agriculture. The dataset that has been pulled includes all studies published on insects. Why might we be interested in the ecotoxicology of neonicotinoids on insects? Feel free to do a brief internet search if you feel you need more background information.

> Answer: We might be interested because the neonicotinoids insecticides especially to bees are the most economically essential and prominent group of pollinators recently in the world. It has a significant solution for crop protection against piercing-sucking pests, and a highly effective way in controlling flea on dogs and cats.

3.  The Niwot Ridge litter and woody debris dataset was collected from the National Ecological Observatory Network, which collectively includes 81 aquatic and terrestrial sites across 20 ecoclimatic domains. 32 of these sites sample forest litter and woody debris, and we will focus on the Niwot Ridge long-term ecological research (LTER) station in Colorado. Why might we be interested in studying litter and woody debris that falls to the ground in forests? Feel free to do a brief internet search if you feel you need more background information.

> Answer: Litter and woody debris become a particular connection between tree canopy and the soils beneath that could influence the forrest productivity and tree growth. It is important to study them sinct it has a significant factor in driving a more sustainable ecological system within the forest. 

4.  How is litter and woody debris sampled as part of the NEON network? Read the NEON_Litterfall_UserGuide.pdf document to learn more. List three pieces of salient information about the sampling methods here:

> Answer: 
 1. Temporal sampling design has a limited access during winter months, this situation may be paused for up to 6 months during the dormant season.
 2. Spatial sampling design has strictly provisions to comply with. 
 3. Temporal sampling should be conducted in various site according to what kind of vegetation such as frequent sampling (1x every 2weeks) in deciduous forest sites during senescence, and infrequent year-round sampling (1x every 1-2 months) at evergreen sites.

## Obtain basic summaries of your data (Neonics)

5.  What are the dimensions of the dataset?

```{r}
dim.data.frame(Neonics.data)
dim(Neonics.data)

```

6.  Using the `summary` function on the "Effect" column, determine the most common effects that are studied. Why might these effects specifically be of interest?

```{r}
summary(Neonics.data$Effect)
summary(Neonics.data)
```

> Answer: The most common effects in more than 1000 times are mortality (1493) and population (1803)

7.  Using the `summary` function, determine the six most commonly studied species in the dataset (common name). What do these species have in common, and why might they be of interest over other insects? Feel free to do a brief internet search for more information if needed.

```{r}
summary(Neonics.data$Species.Common.Name)

```

> Answer: Six most commonly studied species are Honey Bee (667), Parasitic Wasp (285), Buff Tailed Bumblebee (183), Carniolan Honey Bee (152), Bumble Bee (140), and Italian Honeybee (113).
What do these species have in common, and why might they be of interest over other insects? 

8.  Concentrations are always a numeric value. What is the class of Conc.1..Author. in the dataset, and why is it not numeric?

```{r}
class(Neonics.data$Conc.1..Author.)
```

> Answer: The class of Conc.1..Author. in the dataset is a factor because it comprises in leveling according to data of author.

## Explore your data graphically (Neonics)

9.  Using `geom_freqpoly`, generate a plot of the number of studies conducted by publication year.

```{r}
ggplot(Neonics.data) + geom_freqpoly(aes(x = Publication.Year), bins = 15)


```

10. Reproduce the same graph but now add a color aesthetic so that different Test.Location are displayed as different colors.

```{r}
ggplot(Neonics.data) + geom_freqpoly(aes(x = Publication.Year), bins = 15) + geom_freqpoly(aes(x = Publication.Year, color = Test.Location), bins = 15)
```

Interpret this graph. What are the most common test locations, and do they differ over time?

> Answer: The most common test location is Lab with an increasing number over the years

11. Create a bar graph of Endpoint counts. What are the two most common end points, and how are they defined? Consult the ECOTOX_CodeAppendix for more information.

```{r}
ggplot(Neonics.data, aes(x = Endpoint)) + geom_bar()
summary(Neonics.data$Endpoint)

```

> Answer: Two most common end point: 1. NOEL (1816 endpoints) is No-observable-effect-level or the highest dose (concentration) producing effects not significantly  different from responses of controls according to author's reported statistical test (NOEAL/NOEC)  and; 2. LOEL or Lowest-observable-effect-level with 1664 endpoint number that describing the lowest dose (concentration) producing effects that were significantly different (as reported by authors) from responses of controls (LOEAL/LOEC) 

## Explore your data (Litter)

12. Determine the class of collectDate. Is it a date? If not, change to a date and confirm the new class of the variable. Using the `unique` function, determine which dates litter was sampled in August 2018.

```{r}
class(Litter.data$collectDate)

Litter.data$collectDate <- as.Date(Litter.data$collectDate, format = "%Y-%m-%d")
class(Litter.data$collectDate)

unique(Litter.data$collectDate)

```

13. Using the `unique` function, determine how many plots were sampled at Niwot Ridge. How is the information obtained from `unique` different from that obtained from `summary`?

```{r}
unique(Litter.data$plotID)
summary(Litter.data$plotID)
```

> Answer: The function of "unique" will shows the plotID and eliminating the data duplication in the vectors, while summary shows all the plot ID without any elimination process. This process of unique function will also work on other ype of data in matrix and dataframe. 

14. Create a bar graph of functionalGroup counts. This shows you what type of litter is collected at the Niwot Ridge sites. Notice that litter types are fairly equally distributed across the Niwot Ridge sites.

```{r}
ggplot(Litter.data, aes(x = functionalGroup)) + geom_bar()

```

15. Using `geom_boxplot` and `geom_violin`, create a boxplot and a violin plot of dryMass by functionalGroup.

```{r}
#
ggplot(Litter.data) + geom_boxplot(aes(x = functionalGroup, y = dryMass))

#
ggplot(Litter.data) + geom_violin(aes(x = functionalGroup, y = dryMass), draw_quantiles = c(0.25, 0.5, 0.75))
```

Why is the boxplot a more effective visualization option than the violin plot in this case?

> Answer: It is more effective because boxplot visualize the data more clearly since the datatype is numerical.

What type(s) of litter tend to have the highest biomass at these sites?

> Answer: The needles has the highest biomass among the other type of litters.
