---
title: "Assignment 5: Data Visualization"
author: "Yosia Theo Napitupulu"
output: pdf_document
geometry: margin=2.54cm
---

```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Visualization 

## Directions
1. Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.

The completed exercise is due on Friday, Oct 14th @ 5:00pm.



## Set up your session 
1. Set up your session. Verify your working directory and load the tidyverse, 
lubridate, & cowplot packages. Upload the NTL-LTER processed data files for  
nutrients and chemistry/physics for Peter and Paul Lakes (use the tidy 
[`NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv`] version) and 
the processed data file for the Niwot Ridge litter dataset (use the
[`NEON_NIWO_Litter_mass_trap_Processed.csv`] version). 

2. Make sure R is reading dates as date format; if not change the format to date.

```{r, message = FALSE, warning = FALSE}
#1 Set up, locating working directory
getwd()

# load packages
library(tidyverse)
library(lubridate)
library(cowplot)

NTL.chem.nutrient.data.PeterPaul <- read.csv("./Data/Processed/NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv", stringsAsFactors = TRUE)

NIWO.litter.data <- read.csv("./Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv", stringsAsFactors = TRUE)


#2 Change format date
NTL.chem.nutrient.data.PeterPaul$sampledate <- as.Date(NTL.chem.nutrient.data.PeterPaul$sampledate, format = "%Y-%m-%d")

NIWO.litter.data$collectDate <- as.Date(NIWO.litter.data$collectDate, format = "%Y-%m-%d")

class(NTL.chem.nutrient.data.PeterPaul$sampledate)
class(NIWO.litter.data$collectDate)
```
## Define your theme
3. Build a theme and set it as your default theme.

```{r, message = FALSE, warning = FALSE}
#3 Set theme
mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best 
practices for data visualization. Ensure your theme, color palettes, axes, 
and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus (`tp_ug`) by phosphate (`po4`), with 
separate aesthetics for Peter and Paul lakes. Add a line of best fit and 
color it black. Adjust your axes to hide extreme values (hint: change the 
limits using `xlim()` and/or `ylim()`).

```{r, fig.height = 3, fig.width = 4, warning = FALSE}
#4 total phosphorus (`tp_ug`) by phosphate (`po4)
nutrient.plot <- 
  ggplot(NTL.chem.nutrient.data.PeterPaul, aes(x = po4, y = tp_ug, 
  color = lakename)) + geom_point() + 
  geom_smooth(method = lm, color = "black") + 
  ylim (0, 150) + xlim(0, 50)
print(nutrient.plot)

```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, (b) TP, and 
(c) TN, with month as the x axis and lake as a color aesthetic. Then, create 
a cowplot that combines the three graphs. Make sure that only one legend is 
present and that graph axes are aligned. 

Tip: R has a build in variable called `month.abb` that returns a list of months;
see <https://r-lang.com/month-abb-in-r-with-example>

```{r}
#5 Plots for temperature, TP and TN
#change month from int to factor
NTL.chem.nutrient.data.PeterPaul$month <- as.factor(NTL.chem.nutrient.data.PeterPaul$month)


# boxplot for temperature
temp.plot <- ggplot(NTL.chem.nutrient.data.PeterPaul, aes(x = month, y = temperature_C)) + 
  geom_boxplot(aes(color = lakename))
print(temp.plot)

# boxplot for TP
TP.plot <- ggplot(NTL.chem.nutrient.data.PeterPaul, aes(x = month, y = tp_ug)) + geom_boxplot(aes(color = lakename))
print(TP.plot)

# boxplot for TN
TN.plot <- ggplot(NTL.chem.nutrient.data.PeterPaul, aes(x = month, y = tn_ug)) + geom_boxplot(aes(color = lakename))
print(TN.plot)

```

Question: What do you observe about the variables of interest over seasons  and between lakes? 

> Answer: The both lakes only have a contrast difference of median temperatures value on August and october, while the other months seems quite similar with Peter Lake has a higher median value only on October.
while regarding the tp_ug and tn_ug, Peter Lake has a higher median value of tp_ug and tn_ug through the year of observation.
Another interesting information is about the number of outlier which both lakes has pretty high data outlier in tp_ug and tn_ug. However, Peter Lake still becoming the lake who has more outlier in both tp_ug and tn_ug.

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying only the 
"Needles" functional group. Plot the dry mass of needle litter by date and 
separate by NLCD class with a color aesthetic. (no need to adjust the name 
of each land use)

7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated into 
three facets rather than separated by color. 

```{r}
#6 Plotting the dry mass of needle litter
ggplot(subset(NIWO.litter.data, functionalGroup == "Needles"), aes(x = collectDate, y = dryMass)) + geom_point(aes(color = nlcdClass))


#7 Plotting the dry mass of needle litter with 3 facets of nlcd class
ggplot(subset(NIWO.litter.data, functionalGroup == "Needles"), aes(x = collectDate, y = dryMass)) + geom_point() + facet_wrap(vars(nlcdClass), nrow = 3)

```
Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: I think the plot 7 would be more effective because it has a more obvious picture and visualization style. The plot 7 has segmenting the variables in each category, then it becomes more easy to measure and compare between three of each class in each year, rather than the coloured plot 6.