---
title: "Assignment 2: Coding Basics"
author: "Yosia Theo Napitupulu"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1. Sequence of numbers from one to 100, increasing by fours. Assign this sequence a name
increasebyfour_sequence <- seq(1, 100, 4) #from, to, by

#2. Mean and median of this sequence
mean_sequence <- mean(increasebyfour_sequence)
median_sequence <- median(increasebyfour_sequence)
mean_sequence
median_sequence


#3. Ask R to determine whether the mean is greater than the median.
mean_sequence > median_sequence

#4. I create a variable name "increasebyfour_sequence" and put the sequence into it, calculate the mean and median, then asking R to determine whether mean is greater than the median by doing a conditional statement

```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}
students_name <- c("AMIR", "BADU", "CHARLES", "DODO") # character
test_scores <- c(78, 48, 89, 86)              # numeric
test_result <- c(TRUE, FALSE, TRUE, TRUE)     # logical

#creating a new data frame
students_test_df <- data.frame("name"=students_name, "scores"=test_scores, "result"=test_result)
students_test_df
```

9.  QUESTION: How is this data frame different from a matrix?

> Answer: It is differ from matrix since it only has one mode and one data dimension, while data frame has two dimensional data with more than one mode of data

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r}
passing_grade <- function(x){
  ifelse (x>50, "TRUE", "FALSE")
}
passing_grade(test_scores)
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: If and else could be used but it would need a longer code for each element of the vector, however, ifelse is more simple which will allows us to do it directly in a line
